Corporate Security Solutions for BYOD:\\ A Novel User-Centric and Self-Adaptive System

### 2015-07-09 Response to reviewers - r2.tex

A more detailed definition and description of the sensors has been included. In addition, a discussion on how the case the reviewer mentions could affect the system has been added, as:

\textit{It is important to note that MUSES does not work as an antivirus itself, but has a sensor which checks if the device has already an antivirus, and raises a security violation alert if not. It might happen then that the user ignores this advice and the device becomes compromised. As it would be explained in the next section, the sensors are constantly sending their monitored values, and therefore patterns are created, stored, and processed on the server. In case the device gets infected with rootkit-type malware \cite{bickford2010rootkits}, a sudden change in the values of the sensors will mean a significant change in the patterns. Then, this change in the patterns will be noticed by the server, which will immediately warn the user, and decrease the device trust value.}

\begin{verbatim}
4.      Paper structure is really confusing. Many essential concepts 
are introduced and explained too late in the paper forcing the reader 
to go back and forth to understand the architecture (e.g. MUSES aware 
apps and Data Mining techniques). I suggest authors to deeply revise 
the presentation of their work.
\end{verbatim}

The paper structure has been completely redefined, with an introduction
that presents the work, a section that introduces preliminary concepts
followed by the state of the art in corporate mobile security and,
eventually, the description of the framework that is the main point of
the paper. Changes have been too extensive to include here, but great
care has been taken to talk about concepts only when defined
previously. For instance, the concept of MUSES-Aware apps is now
introduced in the context of the API in Section 3.5 and only used
later on, from Section 3.6.4, in the context of other tool. The
concept of data mining is presented along with machine learning in section 4.3.1. 
We have also added some tables (in Section 5) as an additional summary of information, for the sake of clarity.

@@

\begin{verbatim}
5.      A discussion on research activities on BYOD is completely 
missing. This is unacceptable for a solution that aims at solving 
BYOD security issues. Just to cite a few:
a.      Gessner, Dennis, et al. "Towards a User-Friendly 
Security-Enhancing BYOD Solution." NEC Technical Journal 
7.3 (2013): 113.
b.      Samaras, Vasileios. A BYOD Enterprise Security Architecture 
for accessing SaaS cloud services. Diss. TU Delft, Delft 
University of Technology, 2013.
c.      Armando, Alessandro, et al. "Enabling BYOD through secure 
meta-market." Proceedings of the 2014 ACM conference on Security 
and privacy in wireless and mobile networks. ACM, 2014.
\end{verbatim}

The reviewer is right. Although several works have been cited in other sections of the work, a proper discussion on research on BYOD was necessary. Therefore, we have analysed the papers suggested by the reviewer, along with several new relevant ones. These new references also have been also cited in other sections of the work to justify MUSES features; for example, the set of best practices for BYOD security proposed by H. Romer. These works have been also considered to propose the new taxonomy included in the paper (Section X, Page XX). These references are \cite{Garba15organisational,Scarfo12survey,Gessner13userfriendly,Armando14metamarket,Haejung12Door,Samaras13SaaS,Miller12Privacy,Romer14BestPractices}, and can be found at the end of this document.

\begin{verbatim}
6.      The experimental section is completely missing. Is MUSES 
able to identify security violations? Which are the computational 
cost for both the company (server) and users (mobile devices)? Are 
there any limitations?
\end{verbatim}

We agree with what the reviewer remarks, and so we have created a new Section (4.4, in page XX), describing the trials that have been conducted with the first prototype of MUSES. In that section there are commented the ratio of security violations detected and how the system has influenced in the complete reduction of this number to 0.
This has been also commented in the Introduction of the paper.

The main limitation of MUSES, also commented in Section 4.4, is the need of an initial set of defined security rules to start working. The other limitation comes from the set of deployed sensors, which are the only source of data for the system, so it cannot manage security violations happened by different ways.

To avoid this in part, there is a set of guidelines for the companies to build adequate security policies. For instance, the set of initially defined rules used for the trial included a blacklist of forbidden applications, a whitelist of allowed ones, and requirement of antivirus, among others.

\textcolor{blue}{TODO - computational costs? Al menos en el cliente, y la correlaciÛn + risk and trust en el servidor, ya que no se ha llegado a aplicar lo del KRS. Aunque sea muy aproximado (segundos, milisegundos)}

@@

1) Explain why particular solutions are chosen for inclusion in review 
and what is their relevance? Did authors consider other (possible) 
solutions relying on DM and CI?
\end{verbatim}

\textcolor{blue}{TODO}

\begin{verbatim}
 2) Considering variety of BYOD tools and their different features, 
authors might consider to introduce or reuse a taxonomy or 
classification of those approaches, and state the place of MUSES 
there (or define a new subcategory, if deemed necessary)
\end{verbatim}

Following the reviewer's suggestion we have introduced a taxonomy for the analyzed systems, according to the works that we have revised in the state of the art (Section 2). This taxonomy is described at the beginning of Section 3 (second paragraph, page XX). All the tools have been included in the types defined by the taxonomy.
% Antonio - completar con el n˙mero de p·gina final
% Antonio - TODO: decir si MUSES y el resto de sistemas se han encuadrado dentro de esa taxonomÌa

### 2015-07-09 MUSES_vs_BYOD_solutions.tex
Cambios en el paper y en las revisiones
Con comentarios

On the device side, one of the most powerful solutions found in the BYOD state of the art is, apparently, to directly use a phone which has been developed with data security in mind such as the BlackPhone \cite{Blackphone}. It has its own Android-based operating system, called \textit{PrivatOS} \cite{Blackphone}, which includes a privacy-focused application store, called \textit{Silent Store}, that takes care of the problem of applications which ask for certain permissions that can lead, for instance, to personal data leakage \cite{gangula2013survey}. BlackPhone also allows a remote wiping of the data if the device is lost or stolen. The main disadvantage of this solution is either the enterprise having to make an investment and buy these smartphones to the employees or to make employees buy them, so they cannot use the device they already have. Both options are against the BYOD philosophy. On the contrary, the MUSES system proposed in this work is designed as device-independent.
Furthermore, the lack of information about this system makes difficult to determine its real qualities.

@@

\begin{description}
  \item[Connectivity sensor:] It gathers all information related to network communications, in addition to security properties like network encryption, or number of neighbours. This allows MUSES actuator to forbid the user for sending important emails, or opening key assets if, for instance, the network encryption is not secure enough. 
  \item[Device protection sensor:] Which scans device security settings, such as the time until the device screen turns off, if the device has the screen protected with a password, or if the device is rooted \cite{hildenbrand2014}. This security mechanism adds the possibility of detecting security violations even if the user is not doing a specific action.
  \item[Location sensor:] User location can be obtained using the device GPS. Although, this might be taken against the preservation of user privacy. This is why MUSES allows the definition of a number of `zones', and therefore it only checks if the user is inside them or not, acting in consequence.
  \item[Mail sensor:] If the user is writing an email, MUSES client sensor can also check the email attributes. Thus, it can detect if the event context is dangerous, according to security policies and the RT2AE, before the mail is sent.
  % (Paloma) Pedro, aquí he cambiado el "is going to send" a "is writing" en vez de "is sending" porque en realidad lo que queremos es que se le avise antes de mandarlo, no cuando lo está mandando o lo ha mandado. Se explica después. ¿Cómo lo veis?
  \item[File sensor:] The type of file, also called \textit{asset}, is also very important for the application of security rules. Thus, this sensor aims to detect if some files might need more strict security policies, also depending on the rest of sensors.
  \item[App sensor:] It is in charge of monitoring the information about the application from which the user is performing the action, but also looks for installed trusted antivirus, and encryption applications. Additionally, this sensor gathers information about the applications which the user tries to install or uninstall.
\end{description}

@@

As results, not only the number of security violations were reduced by half in the first two weeks (10977), but the number went on decreasing until it was 0 even before the trials period finished. Thus, we can assume that MUSES is an effective tool, even though it has the aforementioned limitation of needing an initial set of defined rules to start working. For this, MUSES also includes a set of guidelines for the companies to build adequate security policies. For instance, the set of initially defined rules used for the trial included a blacklist of forbidden applications, a whitelist of allowed ones, and requirement of antivirus, among others.

@@

\begin{table}[ht]
\begin{footnotesize}
  \centering
  \renewcommand{\arraystretch}{1.8}% Spread rows out...
%  \resizebox{8cm}{!}{
  \begin{tabular}{>{\centering\bfseries}m{1in} >{\centering}m{1in} >{\centering}m{1in} >{\centering\arraybackslash}m{1in}}
       \toprule
      & \textbf{License} & \textbf{Devices} & \textbf{Price} \\
       \midrule
    MUSES & OpenSource & Android, Windows 8.1, Windows Phone, iOS & Free  \\
       \arrayrulecolor[gray]{0.8}\hline
    IBM Mobile Security & Proprietary & IOS, Android, Windows Phone, Blackberry  &  By Request\\\\
        \arrayrulecolor[gray]{0.8}\hline
    Sophos Mobile Control & Proprietary & IOS, Android, Windows Phone  &  By Request\\
        \arrayrulecolor[gray]{0.8}\hline
    Samsung Knox and Blackbery Balance & Proprietary & Android Samsung Devices and IOS & 12\$ per year/device \\
        \arrayrulecolor[gray]{0.8}\hline
    Blackphone & Proprietary & Own device & 120\$ per year \\
        \arrayrulecolor[gray]{0.8}\hline
    Android for Work & Proprietary & Android and devices with Chrome Browser &  12\$ per month/user\\
       \bottomrule
  \end{tabular}
%  }
\end{footnotesize}
  \caption{Features of the analyzed products considering licenses, type of devices and OS, and price.}
  \label{tab:features}
\end{table}
% Antonio - Antares, ¿no había más tablas?
% Por favor, que alguien rellene los datos de la solución de IBM. ¿Fergu?¿Antares?
% Quizá habría que indicar que MUSES se puede instalar en PC también en la tabla. ¿Sabéis si los demás se puede o son sólo para smartphones/tablets?

>>>>

On the device side, one of the most powerful solutions found in the BYOD state of the art is, apparently, to directly use a phone which has been developed with data security in mind such as the BlackPhone \cite{Blackphone}. It has its own Android-based operating system, called \textit{PrivatOS} \cite{Blackphone}, which includes a privacy-focused application store, called \textit{Silent Store}, that takes care of the problem of applications which ask for certain permissions that can lead, for instance, to personal data leakage \cite{gangula2013survey}. BlackPhone also allows a remote wiping of the data if the device is lost or stolen. The main disadvantage of this solution is either the enterprise having to make an investment and buy these smartphones to the employees or to make employees buy them, so they cannot use the device they already have. Both options are against the BYOD philosophy. On the contrary, the MUSES system proposed in this work is designed as device-independent.
Furthermore, the lack of information about this system makes it difficult to determine its real qualities.

@@

% Antonio - pongo esto como lista de items, ya que era la única lista diferente
\begin{itemize}
  \item \textit{Connectivity sensor:} It gathers all information related to network communications, in addition to security properties like network encryption, or number of neighbours. This allows MUSES actuator to forbid the user for sending important emails, or opening key assets if, for instance, the network encryption is not secure enough. 
  \item \textit{Device protection sensor:} Which scans device security settings, such as the time until the device screen turns off, if the device has the screen protected with a password, or if the device is rooted \cite{hildenbrand2014}. This security mechanism adds the possibility of detecting security violations even if the user is not doing a specific action.
  \item \textit{Location sensor:} User location can be obtained using the device GPS. Although, this might be taken against the preservation of user privacy. This is why MUSES allows the definition of a number of `zones', and therefore it only checks if the user is inside them or not, acting in consequence.
  \item \textit{Mail sensor:} If the user is writing an email, MUSES client sensor can also check the email attributes. Thus, it can detect if the event context is dangerous, according to security policies and the RT2AE, before the mail is sent.
  % (Paloma) Pedro, aquí he cambiado el "is going to send" a "is writing" en vez de "is sending" porque en realidad lo que queremos es que se le avise antes de mandarlo, no cuando lo está mandando o lo ha mandado. Se explica después. ¿Cómo lo veis?
  \item \textit{File sensor:} The type of file, also called \textit{asset}, is also very important for the application of security rules. Thus, this sensor aims to detect if some files might need more strict security policies, also depending on the rest of sensors.
  \item \textit{App sensor:} It is in charge of monitoring the information about the application from which the user is performing the action, but also looks for installed trusted antivirus, and encryption applications. Additionally, this sensor gathers information about the applications which the user tries to install or uninstall.
\end{itemize}

@@

As results, not only the number of security violations were reduced by half in the first two weeks (10977), but the number went on decreasing until it was 0 even before the trials period finished. Thus, we can assume that MUSES is an effective tool, even though it has the aforementioned limitation of needing an initial set of defined security rules to start working. For this, MUSES also includes a set of guidelines for the companies to build adequate security policies. For instance, the set of initially defined rules used for the trial included a blacklist of forbidden applications, a whitelist of allowed ones, and requirement of antivirus, among others.

@@

\begin{table}[ht]
\begin{footnotesize}
  \centering
  \renewcommand{\arraystretch}{1.8}% Spread rows out...
%  \resizebox{8cm}{!}{
  \begin{tabular}{>{\centering\bfseries}m{1in} >{\centering}m{1in} >{\centering}m{1in} >{\centering\arraybackslash}m{1in}}
       \toprule
      & \textbf{License} & \textbf{Devices} & \textbf{Price} \\
       \midrule
    MUSES & OpenSource & Android, Windows 8.1, Windows Phone, iOS & Free  \\
       \arrayrulecolor[gray]{0.8}\hline
    IBM Mobile Security & Proprietary & iOS, Android, Windows Phone, Blackberry  &  By Request\\\\
        \arrayrulecolor[gray]{0.8}\hline
    Sophos Mobile Control & Proprietary & iOS, Android, Windows Phone  &  By Request\\
        \arrayrulecolor[gray]{0.8}\hline
    Samsung Knox and Blackbery Balance & Proprietary & Android Samsung Devices and iOS & 12\$ per year/device \\
        \arrayrulecolor[gray]{0.8}\hline
    Blackphone & Proprietary & Own device & 120\$ per year \\
        \arrayrulecolor[gray]{0.8}\hline
    Android for Work & Proprietary & Android and devices with Chrome Browser &  12\$ per month/user\\
       \bottomrule
  \end{tabular}
%  }
\end{footnotesize}
  \caption{Features of the analyzed products considering licenses, type of devices and OS, and price.}
  \label{tab:features}
\end{table}
% Antonio - Antares, ¿no había más tablas?
% Por favor, que alguien rellene los datos de la solución de IBM. ¿Fergu?¿Antares?
% Quizá habría que indicar que MUSES se puede instalar en PC también en la tabla. ¿Sabéis si los demás se puede o son sólo para smartphones/tablets?
